\paperTitle

### 2014-05-13 doc/icml2014/directed.tex
Updated exclusive views property to be true with stronger assumptions

% DONE: don't talk about parameter recovery yet, just establish exclusive views
%Finally, the claim we make is that we can recover parameters for any
  %graphical model where every hidden variable is a bottleneck:
Now define the following property:
\begin{property}[Uniformly bottlenecked]
  \label{prop:bottleneck}
  A clique $\sC$ is \emph{uniformly bottlenecked} if
  every hidden variable in $\sC$ is a bottleneck. 
  If every clique $\sC \in \sG$ is uniformly bottlenecked, then we say
  that the graph $\sG$ is uniformly bottlenecked as well.
\end{property}
Note that two hidden variables are allowed to have overlapping views.
%If every hidden variable is a bottleneck, then we can estimate $\mOpp{v}{i}$
It is clear that having uniform bottlenecks implies part (ii)
of the exclusive views property (\propertyref{exclusive-views}) via \TensorFactorize,
but we will show that part (i) follows as well, given basic assumptions
on the conditional independence structure of the graphical model.

\begin{assumption}[Conditional Independence]
  \label{asm:ci}  
  \
  \begin{enumerate}
    \item If, conditioned on a variable $h$, two variables $x_1, x_2$
      are independent, then $x_1$ and $x_2$ are not both dependent on
      any other variable $h'$: $x_1 \perp x_2 \given h \implies x_1
      \perp h' \given h \vee x_2 \perp h' \given h$ for every $h' \in
      \sC$.
    \item
      For every clique $\sC \in \sG$, conditioned on a variable $h \in
      \sC$, all other variables in the clique are dependent:
      $h_i \not\perp h_j \given h$ for every $h_i, h_j \in \sC$.
    \item
      Given an active path in the unconditioned model, then it remains
        active when conditioned on $h$. 
      Let $a, b, c$ be three variables in the graph such that
      $$ a \not\perp b \wedge b \not\perp c \implies a \not\perp c,$$
      then,
      $$ a \not\perp b \given h \wedge b \not\perp c \given h \implies a \not\perp c \given h.$$
%      Suppose $$x \not\perp h_1 \wedge h_1 \not\perp h_2 \implies
%      x \not\perp h_2,$$ then $$x \not\perp h_1 \given h \wedge h_1 \not\perp h_2
%      \given h \implies x \not\perp h_2 \given h$$.
  \end{enumerate}
\end{assumption}

%TODO: Describe with example the cliques ruled out.

@@

Now we establish the link between bottlenecks and exclusive views
(see \appendixref{exclusive-views} for the complete proof):
\begin{lemma}[Uniformly bottlenecked implies exclusive views]
  \label{lem:bottleneck-views}  
% PL: simplify to cliques (this is also bad notation, since \sG is a set of cliques, not nodes)
%Let $\sG' = \{ h_{i_1}, \dots, h_{i_m} \} \subseteq \sG$ be
  %a connected subgraph of hidden variables.
  %If every hidden variable
  %$h_i \in \sG'$ is a bottleneck, then $\sG'$ has the exclusive views property.
  If a graphical model $\sG$ satisfies conditional independence
  assumptions (\assumptionref{ci}) and is uniformly bottlenecked, then
  $\sG$ has the exclusive views property.
\end{lemma}
% PL: save space, doesn't add much
%\begin{proof}
%  The independence conditions required by the bottleneck condition
%  effectively guarantee that at least one of the three views for
%  a bottleneck be exclusive. 
%  \appendixref{exclusive-views} provides a complete proof.
%\end{proof}

>>>>

% DONE: don't talk about parameter recovery yet, just establish exclusive views
%Finally, the claim we make is that we can recover parameters for any
  %graphical model where every hidden variable is a bottleneck:
Now define the following property:
\begin{property}[Uniformly bottlenecked]
  \label{prop:bottleneck}
  A clique $\sC$ is \emph{uniformly bottlenecked} if
  every hidden variable in $\sC$ is a bottleneck. 
  If every clique $\sC \in \sG$ is uniformly bottlenecked, then we say
  that the graph $\sG$ is uniformly bottlenecked as well.
\end{property}
Note that two hidden variables are allowed to have overlapping views.
%If every hidden variable is a bottleneck, then we can estimate $\mOpp{v}{i}$
It is clear that having uniform bottlenecks implies part (ii)
of the exclusive views property (\propertyref{exclusive-views}) via \TensorFactorize,
but we will show that part (i) follows as well, given that the graphical model faithfully represents conditional independences:

\begin{assumption}[Faithful Representation]
  \label{asm:ci}  
  If there is an active trail from any two variables $a, b$ in the graph
    $\sG$, then they are not independent, i.e $a \not\perp b$. 
  If there is an active trail between $a$ and $b$ conditioned on any variable,
    say $c$, then they are also not independent, i.e. $a\not\perp b \given c$.
\end{assumption}

@@

Now we establish the link between bottlenecks and exclusive views
(see \appendixref{exclusive-views} for the complete proof):
\begin{lemma}[Uniformly bottlenecked implies exclusive views]
  \label{lem:bottleneck-views}  
% PL: simplify to cliques (this is also bad notation, since \sG is a set of cliques, not nodes)
%Let $\sG' = \{ h_{i_1}, \dots, h_{i_m} \} \subseteq \sG$ be
  %a connected subgraph of hidden variables.
  %If every hidden variable
  %$h_i \in \sG'$ is a bottleneck, then $\sG'$ has the exclusive views property.
  If a graphical model $\sG$ satisfies the faithful representation 
  assumption (\assumptionref{ci}) and is uniformly bottlenecked, then
  $\sG$ has the exclusive views property.
\end{lemma}
% PL: save space, doesn't add much
%\begin{proof}
%  The independence conditions required by the bottleneck condition
%  effectively guarantee that at least one of the three views for
%  a bottleneck be exclusive. 
%  \appendixref{exclusive-views} provides a complete proof.
%\end{proof}

### 2014-05-13 doc/icml2014/exclusive-views.tex
Updated exclusive views property to be true with stronger assumptions

In \sectionref{directed}, we stated that bottlenecks guaranteed
exclusive views provided conditional independence assumptions \assumptionref{ci}. 
We provide the proof below.
  We generalize the lemma so that it applies
  on any connected subgraph, not just cliques.
  Also, the statement is made in terms of subgraphs,
  not the entire graph.

\begin{lemma*}[Bottlenecks guarantee exclusive views]
  %\label{lem:bottleneck-views}  
  Let $\sG$ be a graphical model satisfying conditional independence
  assumptions (\assumptionref{ci}) and be uniformly bottlenecked, then
  $\sG$ has the exclusive views property.
%Let $\sG' = \{ h_{i_1}, \ldots, h_{i_m} \} \subseteq \sG$ be
%  a connected subgraph of hidden variables. If every hidden variable
%  $h_i \in \sG'$ is a bottleneck, then $\sG'$ has the exclusive view property.
\end{lemma*}

\begin{proof}
Let $h_0$ be a hidden variable in the clique $\sC$ and $x_1, x_2$ be any
  two conditionally independent observed variables. 
One of $x_1$ or $x_2$ is an exclusive view for $h_0$ iff it is
  independent of all other hidden variables in the clique, given $h_0$.
Intuitively, if this weren't the case, then we would be able to find
  an active path between $x_1$ and $x_2$ that does not go through $h_0$,
  violating the conditional independence condition.

% TODO: One line here about how assumptions save us.
% The conditional independence assumptions (\assumptionref{ci}) exclude
%   edge cases wherein the active trail still preserves 
% for them to be both depend on any other
% variable $h'$ in the clique.

Suppose neither $x_1$ nor $x_2$ were exclusive views for $h_0$, 
  then there exists some $h_1, h_2 \in \sC$ (not necessarily distinct)  that $x_1, x_2$ are
  respectively dependent on:
  $x_1 \not\perp h_1 \given h_0$ and $x_2 \not\perp h_2 \given h_0$. 
Both $h_1$ and $h_2$ are part of the same clique $\sC$, thus by
  \assumptionref{ci}, we have that $h_1 \not\perp h_2 \given h_0$.

Since $x_1 \not\perp h_1 \given h$, there exists an active trail from
$x_1 \to h_1$. Furthermore, because $h_1$ and $h_2$ are in the same
clique, the active trail extends to $h_2$, i.e. $x_1 \not\perp h_2$.
However, since $x_1 \to h_1 \to h_2$ is an active trail in the unconditioned graph, i.e. $x_1 \not\perp h_1$; by \assumptionref{ci}, it remains active trail when conditioned on $h$.

Consequently, $x_1 \not\perp h_2 \given h_0$, which
contradicts the assumption that as $x_1 \perp x_2 \given h_0$,
$x_1 \not\perp h_i \given h_0$ and $x_2 \not\perp h_i \given h_0$ for all $h_i \in \sC$.

>>>>

In \sectionref{directed}, we stated that bottlenecks guaranteed
exclusive views provided the graph faithfully represents conditional independences (\assumptionref{ci}). 
We provide the proof below.
  We generalize the lemma so that it applies
  on any connected subgraph, not just cliques.
  Also, the statement is made in terms of subgraphs,
  not the entire graph.

\begin{lemma*}[Bottlenecks guarantee exclusive views]
  %\label{lem:bottleneck-views}  
  Let $\sG$ be a graphical model satisfying the faithful representation assumption (\assumptionref{ci}).
  Let $\sC \in \sG$ be a uniformly bottlenecked clique, then
  $\sC$ has the exclusive views property, i.e.\ every variable in $\sC$
  has an exclusive view.
%Let $\sG' = \{ h_{i_1}, \ldots, h_{i_m} \} \subseteq \sG$ be
%  a connected subgraph of hidden variables. If every hidden variable
%  $h_i \in \sG'$ is a bottleneck, then $\sG'$ has the exclusive view property.
\end{lemma*}

\begin{proof}
By the canonical reduction (\lemmaref{reduction}), we can assume without
  loss of generality that all observed variables are leaves and thus
  all v-structures are active.
Combined with the faithful representation assumption, this lets us
  reason about independences solely in terms of graph connectivity.

Let $h_0$ be a hidden variable in the clique $\sC$ and let $x_1, x_2$ be any
  two conditionally independent observed variables. 
Recall that if $x_1$ is an exclusive view of $h_0$ in $\sC$,
  then, conditioned on $h_0$, $x_1$ is independent of every other variable $h_i
  \in \sC$.
% Since $x_1 \perp x_2 \given h_0$, every path from $x_1$ to $x_2$ must go through $h_0$. 
Suppose neither $x_1$ nor $x_2$ are exclusive views for $h_0$, 
  then there exists some $h_1, h_2 \in \sC$ that $x_1, x_2$ are respectively dependent on.
However, $h_1$ and $h_2$ are connected by participation in the clique; consequently there is a path $x_1 \to h_1 \to h_2 \to x_2$. 
By \assumptionref{ci}, this implies that $x_1$ and $x_2$ are not
  independent conditioned on $h_0$, which is a contradiction.

Thus, at least one of $x_1$ or $x_2$ is isolated from the rest of the
variables in the clique and forms an exclusive view for $h_0$ in $\sC$.

### 2014-05-13 doc/icml2014/undirected.tex
Updated exclusive views property to be true with stronger assumptions

The pseudolikelihood also replaces the log-partition function $A(\theta)$
with a sum over the variables in the neighbourhood: $A_a(\theta; x_{\sN(a)}) = \sum_{a' \in \{a\} \union \sN(a)} \exp(\theta^\top\phi_{\sN(a)}(a'))$,
which cannot be computed from just the hidden clique marginals $Z_\sC$,
since the neighbors $\sN(a)$ in general do not form a clique.
We adapt our earlier algorithm ($\LearnClique$) to estimate the
  marginals over all neighborhoods $Z_{\sN(a)}$ by noting that
  \lemmaref{bottleneck-views} guarantees exclusive views for any connected
  subgraph of $\sG$, not just cliques.

>>>>

The pseudolikelihood also replaces the log-partition function
  $A(\theta)$ with a sum over the variables in the neighborhood:
  $A_a(\theta; x_{\sN(a)}) = \sum_{a' \in \{a\} \union \sN(a)}
  \exp(\theta^\top\phi_{\sN(a)}(a'))$.
Note that the neighbors $\sN(a)$ in general do not form a clique, and
  thus the uniformly bottlenecked assumption alone is not sufficient to
  guarantee exclusive views for each variable in the subgraph.
However, when we do have exclusive views for each neighborhood, as in
  \figureref{examples-mrf}, we can adapt our earlier algorithm
  ($\LearnClique$) to estimate the marginals over all neighborhoods $Z_{\sN(a)}$.

% Arun: WRONG
% The pseudolikelihood also replaces the log-partition function $A(\theta)$
% with a sum over the variables in the neighborhood: $A_a(\theta; x_{\sN(a)}) = \sum_{a' \in \{a\} \union \sN(a)} \exp(\theta^\top\phi_{\sN(a)}(a'))$,
% which cannot be computed from just the hidden clique marginals $Z_\sC$,
% since the neighbors $\sN(a)$ in general do not form a clique.
% We adapt our earlier algorithm ($\LearnClique$) to estimate the
%   marginals over all neighborhoods $Z_{\sN(a)}$ by noting that
%   \lemmaref{bottleneck-views} guarantees exclusive views for any connected
%   subgraph of $\sG$, not just cliques.
